Urban Heat Risk Mapping Using Multiple Point Patterns in Houston, Texas

### 2016-11-02 jrssc-mpp.tex
Added black and white plots, work on results section

@@

\begin{figure}
  \centering
  \begin{minipage}[t]{1.0\textwidth}
    \centering
    \includegraphics[width=0.7\textwidth]{./imgs/911_kde.png}
    \\
    a) 911 Data
    \\
  \end{minipage}
  \vfill
  \begin{minipage}[t]{1.0\textwidth}
    \centering
    \includegraphics[width=0.7\textwidth]{./imgs/mortality_kde.png}
    \\
    b) Mortality Data
    \\
  \end{minipage}
  \caption{Kernel Density Estimates for locations of heat-related 911 calls and mortalities. 911 call data courtesy of the Houston Fire Department. Mortality data courtesy of the Texas Department of State Health Services.}
  \label{kde}
\end{figure}

Consider, for example, the two sources of data displayed in 
Figure \ref{kde}. Figure \ref{kde}a displays a kernel density estimate 
of the locations to which ambulances were dispatched in 2054 heat-related emergency calls to 911 occuring between
May and September 2006-2010 classified under at least one of the
terms ``heat,'' ``dehydration,'' ``heat exhaustion,'' or ``heat
stroke'' (exact locations of the calls cannot be shown in order
to preserve confidentiality). Figure \ref{kde}b, however, displays
a kernel density estimate of the residences of 15,244 mortalities
that occurred between May and September of the years 1999 through
2006 where the primary cause of death was connected to either
cardiovascular or kidney problems. Inevitably, some of the deaths in this group will not be directly heat-related, but both causes
have been shown to have a strong relationship with heat \citep{Kjellstrom2010}.

Examination of the kernel density estimates reveals that while the 
density of 911 calls is concentrated on a single location in the 
center of Houston, the density for the mortality data is more
diffuse, exhibiting three distinct areas of concentration. 
The fact that the spatial distribution of heat-related 911 calls
is different than that of heat-related mortalities suggests that
each measure captures only part of the overall effects of heat on
health. As a result, being able to combine the information
contained in each of the different data sources into a single risk
map in a systematic, model-based fashion becomes a useful tool
in understanding how different areas in a city are impacted 
by extreme heat as opposed to a more narrow view of where a
911 call or mortality is most likely to take place.

@@

\subsection{Model Fitting}
\textcolor{red}{We fit this model within a Bayesian framework. For each health outcome $p = 1, \dots, P$, randomly generate draws for $\delta_p$ from the known posterior, $\mathcal{G}(n_p + 0.001, 1.001)$. Similarly, because the priors
for $\boldsymbol{\alpha}_p$, $\boldsymbol{\gamma}_p$, and $\boldsymbol{\xi}_p$ are all conjugate, their posterior
distributions are known and can be drawn from directly. That is, we can easily obtain random draws from $\text{Dir}(n_{p,0} + 1, n_{p,1} + 1, \dots, n_{p,113})$, $\text{Dir}(n_{p,Female} + 1, n_{p,Male} + 1)$, and $\text{Dir}(n_{p,Asian} + 1, \dots, n_{p,White} + 1)$, where as before $n_{p,0}$ denotes the number of people who experienced health outcome $p$ at age $0$, $n_{p,Female}$ indicates the number of females who experienced health outcome $p$, and so on.}

@@

Note that for steps \ref{draws2}, \ref{drawsmu}, and \ref{drawbeta} we have the complete conditional distributions, but that \ref{drawphi} and \ref{drawmu} require a Metropolis-Hastings (MH) update. Using a sampler of this design, 10000 draws were taken from the posterior distributions for each of these parameters, after a burn-in of 10000. Monte Carlo standard error (MCSE) was calculated for each chain and provided evidence that all of them converged \citep{Flegal2015}.

@@

\subsection{Covariates and Variable Selection}\label{sec:temp_data}
Recall that in \eqref{mu} we regressed $\boldsymbol{\mu}$ on covariates $\mathbf{X}$ in order to account for potential confounding effects within each grid cell.   For this analysis, the $\mathbf{X}$ matrix denoted above consists of an intercept, temperature data defined below, the percent of people over 65 in each grid cell and the number of people without central air conditioning in a given grid cell. Percent of people over 65 was taken from 2010 U.S. Census Data and the number of people without air conditioning was taken from the Harris County tax assessor parcel database. Both were aligned with our discrete grid using the method described in Section \ref{likelayer}.

In order to get heat information at the resolution necessary for this
study, we used temperature data simulated using an offline version of the 
Noah Land Surface Model, known as the High Resolution Land Data Assimilation 
System (HRLDAS).
These data consist of 15,625 HRLDAS 1-km$^2$ grid boxes that cover the Houston
area and contain 6 temperature measurements that we consider: daily max/min 
temperature at 2 meters above the ground (T2), daily max/min NWS heat index (HI), and
daily max/min Swedish wet-bulb-globe temperature (SWBG). To get a measure of the 
``heat'' associated with locations across the city, the values for each of 
these variables were averaged across the months May through September of 
1999 -- 2010 (the time window for all observed health outcomes).

We fit the model in Chapter \ref{model} using each of these temperature variables (one at a time) and then calculated the deviance information criterion (DIC) \citep{Spiegelhalter2002} as a measure of model fit. 
Table 1%\ref{DIC} 
shows that the model produced the lowest DIC value when fit with max HI so the numerical estimates and risk maps shown in this chapter are from the model fit with this temperature measurement. Figure \ref{fig:temp} displays the average maximum HI values associated with each grid cell and shows a clear spatial distribution, with higher temperatures concentrated near the city center. Note that the DIC values in Table 1
%\ref{DIC} 
are all very close, indicating little difference in model fit between the different variables.  This is, perhaps, due to the fact that the covariates enter our model at the mean intensity surface layer (in the regression for $\boldsymbol{\mu}$) at which point it is difficult to distinguish between the different heat variables. That is, because $\boldsymbol{\mu}$ captures both the 911 calls and mortalities, it is possible that any of the measured heat variables explain the combined impact of heat on ``health'' whereas specific heat variables better explain specific health outcomes.

% Table created by stargazer v.5.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Tue, Nov 03, 2015 - 15:11:28
\begin{table}%\centering 
  \caption{DIC Values for Models Fit Using Each Temperature Measurement.} 
  \label{DIC} 
\begin{tabular}{@{\extracolsep{5pt}} cc} 
\\[-1.8ex]\hline
\hline \\[-1.8ex]  
Temperature Measurement & DIC \\
\hline \\[-1.8ex]  
HI Max & $220991.6$ \\ 
HI Min & $221016.2$ \\ 
T2 Max & $220998.3$ \\ 
T2 Min & $220999.9$ \\ 
SWBG Max & $221008.0$ \\ 
SWBG Min & $221021.4$ \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}

@@

>>>>

\newcommand*\rot{\rotatebox{90}}

@@

\begin{figure}
  \centering
  \begin{minipage}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{./imgs/911_contours.pdf}
    \\
    a) 911 Data
    \\
  \end{minipage}
  \hfill
  \begin{minipage}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{./imgs/mortality_contours.pdf}
    \\
    b) Mortality Data
    \\
  \end{minipage}
  \caption{Contour plots for kernel density estimates of the locations of heat-related 911 calls and mortalities. 911 call data courtesy of the Houston Fire Department. Mortality data courtesy of the Texas Department of State Health Services.}
  \label{kde}
\end{figure}

Consider, for example, the two sources of data displayed in 
Figure \ref{kde}. Figure \ref{kde}a displays a contour plot 
of a kernel density estimate of the locations to which ambulances were dispatched in 2054 heat-related emergency calls to 911 occuring between
May and September 2006-2010 classified under at least one of the
terms ``heat,'' ``dehydration,'' ``heat exhaustion,'' or ``heat
stroke'' (exact locations of the calls cannot be shown in order
to preserve confidentiality). Figure \ref{kde}b, however, displays
a contour plot of the residences of 15,244 mortalities
that occurred between May and September of the years 1999 through
2006 where the primary cause of death was connected to either
cardiovascular or kidney problems. Inevitably, some of the deaths in this group will not be directly heat-related, but both causes
have been shown to have a strong relationship with heat \citep{Kjellstrom2010}.

Examination of the contour plots reveals that while the 
density of 911 calls is concentrated on a single location in the 
center of Houston, the density for the mortality data is more
diffuse, exhibiting three distinct areas of concentration. 
The fact that the spatial distribution of heat-related 911 calls
is different than that of heat-related mortalities suggests that
each measure captures only part of the overall effects of heat on
health. As a result, being able to combine the information
contained in each of the different data sources into a single risk
map in a systematic, model-based fashion becomes a useful tool
in understanding how different areas in a city are impacted 
by extreme heat as opposed to a more narrow view of where a
911 call or mortality is most likely to take place.

@@

\subsection{Model Fitting}
\textcolor{red}{We fit this model within a Bayesian framework. For each health outcome $p = 1, \dots, P$, randomly generate draws for $\delta_p$ from the known posterior, $\mathcal{G}(n_p + 0.001, 1.001)$. Similarly, because the priors
for $\boldsymbol{\alpha}_p$, $\boldsymbol{\gamma}_p$, and $\boldsymbol{\xi}_p$ are all conjugate, their posterior
distributions are known and can be drawn from directly. That is, for $p = 1, \dots, P$ we can easily obtain random draws from $\text{Dir}(n_{p,0} + 1, n_{p,1} + 1, \dots, n_{p,113})$, $\text{Dir}(n_{p,Female} + 1, n_{p,Male} + 1)$, and $\text{Dir}(n_{p,Asian} + 1, \dots, n_{p,White} + 1)$, where as before $n_{p,0}$ denotes the number of people who experienced health outcome $p$ at age $0$, $n_{p,Female}$ indicates the number of females who experienced health outcome $p$, and so on.}

@@

Note that for steps \ref{draws2}, \ref{drawsmu}, and \ref{drawbeta} we have the complete conditional distributions, but that \ref{drawphi} and \ref{drawmu} require a Metropolis-Hastings (MH) update. Using a sampler of this design, 10000 draws were taken from the posterior distributions for each of these parameters, after a burn-in of 10000. In order to assess convergence, we used Monte Carlo standard error (MCSE). MCSE provides evidence of convergence by Monte Carlo standard error (MCSE) was calculated for each chain and provided evidence that all of them converged \citep{Flegal2015}.

@@

\subsection{Covariates and Variable Selection}\label{sec:temp_data}
In order to get heat information at the resolution necessary for this
study, we used temperature data simulated using an offline version of the 
Noah Land Surface Model, known as the High Resolution Land Data Assimilation 
System (HRLDAS).
These data consist of 15,625 HRLDAS 1-km$^2$ grid boxes that cover the Houston
area and contain 6 temperature measurements that we consider: daily max/min 
temperature at 2 meters above the ground (T2), daily max/min NWS heat index (HI), and
daily max/min Swedish wet-bulb-globe temperature (SWBG). To get a measure of the 
``heat'' associated with locations across the city, the values for each of 
these variables were averaged across the months May through September of 
1999 -- 2010 (the time window for all observed health outcomes).

We fit the model in Section \ref{model} using each of these temperature variables (one at a time) and then calculated the deviance information criterion (DIC) \citep{Spiegelhalter2002} as a measure of model fit. 
Table \ref{DIC} 
shows that the model produced the lowest DIC value when fit with max HI so the numerical estimates and risk maps shown in this chapter are from the model fit with this temperature measurement. Figure \ref{fig:temp} displays the average maximum HI values associated with each grid cell and shows a clear spatial distribution, with higher temperatures concentrated near the city center. Note that the DIC values in Table \ref{DIC} 
are all very close, indicating little difference in model fit between the different variables.  This is, perhaps, due to the fact that the covariates enter our model at the mean intensity surface layer (in the regression for $\boldsymbol{\mu}$) at which point it is difficult to distinguish between the different heat variables. That is, because $\boldsymbol{\mu}$ captures both the 911 calls and mortalities, it is possible that any of the measured heat variables explain the combined impact of heat on ``health'' whereas specific heat variables better explain specific health outcomes.

\textcolor{red}{Recall that in \eqref{mu} $\boldsymbol{\mu}$ can be regressed on covariates $\mathbf{X}$ in order to account for potential confounding effects. For all models that we considered $\mathbf{X}$ included an intercept term and temperature data detailed above. Table \ref{tab:model_comparison} shows which additional demographic variables were included in $\mathbf{X}$ as well as indicating whether or not two-way interaction terms were also included in each model. The DIC in the final column of the table indicates that Model X provides the best fit for the data, and so the rest of this paper refers to this model. The percentage of people without air conditioning was taken from the Harris County tax assessor parcel database while all remaining covariates were taken from 2000 U.S. Census Data. Both data sources were aligned with our discrete grid using the method described in Section \ref{likelayer}. }

% Table created by stargazer v.5.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Tue, Nov 03, 2015 - 15:11:28
\begin{table}
  \caption{DIC values for models fit with different temperature measurements.} 
  \label{DIC} 
\begin{tabular}{@{\extracolsep{5pt}} cc} 
\\[-1.8ex]\hline
\hline \\[-1.8ex]  
Temperature Measurement & DIC \\
\hline \\[-1.8ex]  
HI Max & $220991$ \\ 
HI Min & $221016$ \\ 
T2 Max & $220998$ \\ 
T2 Min & $220999$ \\ 
SWBG Max & $221008$ \\ 
SWBG Min & $221021$ \\ 
\hline \\[-1.8ex] 
\end{tabular} 
\end{table}

\begin{table}
  \caption{In order to control for potential confounding effects, we compared 15 different models that included a variety of demographic variables and two-way interactions. The variables included in each model and whether or not interaction terms were considered is indicated in the table below. We also include DIC in order to compare models and select the one that provides the best fit.}
  \label{tab:model_comparison}

  \begin{tabular}{@{\extracolsep{5pt}} lccccccccc|c}
  \\[-1.8ex]\hline
  \hline \\[-1.8ex]
  & \multicolumn{9}{c}{Variables} & \\
  & \rot{\% over 65} & \rot{\% under 5} & \rot{\% Black} & \rot{\% Hispanic} & \rot{\% w/o AC} & \rot{\% disabled} & \rot{\% living alone} & \rot{\% in poverty} & \rot{Interactions} & DIC \\\hline
  Model 01 &   &   &   &   &   &   &   &   &   & 12345 \\
  Model 02 & X & X &   &   &   &   &   &   &   & 12345 \\
  Model 03 &   &   & X & X &   &   &   &   &   & 12345 \\
  Model 04 &   &   &   &   & X & X & X & X &   & 12345 \\
  Model 05 & X & X & X & X &   &   &   &   &   & 12345 \\
  Model 06 & X & X &   &   & X & X & X & X &   & 12345 \\
  Model 07 &   &   & X & X & X & X & X & X &   & 12345 \\
  Model 08 & X & X & X & X & X & X & X & X &   & 12345 \\
  Model 09 & X & X &   &   &   &   &   &   & X & 12345 \\
  Model 10 &   &   & X & X &   &   &   &   & X & 12345 \\
  Model 11 &   &   &   &   & X & X & X & X & X & 12345 \\
  Model 12 & X & X & X & X &   &   &   &   & X & 12345 \\
  Model 13 & X & X &   &   & X & X & X & X & X & 12345 \\
  Model 14 &   &   & X & X & X & X & X & X & X & 12345 \\
  Model 15 & X & X & X & X & X & X & X & X & X & 12345 \\
  \hline\\[-1.8ex]
  \end{tabular}

\end{table}

@@

\begin{figure}
  \centering
  \begin{minipage}{0.49\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{./imgs/delta_911.pdf}
  \end{minipage}
  \hfill
  \begin{minipage}{0.49\textwidth}
    \centering
    \includegraphics[width=1.0\textwidth]{./imgs/delta_mortality.pdf}
  \end{minipage}
  \caption{Posterior density plots for $\delta_{911}$ and $\delta_{Mort}$.}
  \label{kde}
\end{figure}
